Making Deep Q-learning Methods Robust to Time Discretization

### 2019-01-23 paper/icml_drau/experiments.tex

%! TEX root = icml_drau.tex
\section{Experiments}
\label{sec:exp}
\input{lc_fig.tex}
The experiment set provided here is specifically aimed at showing that
the proposed method DAU can operate efficiently on a wide range of time
discretizations, without specific tuning, while standard deep Q-learning
approaches cannot. DAU is compared to DDPG in continuous action settings and to DQN in
discrete action settings.
Since our focus is the study of Q-learning based off-policy methods, we do not
compare against on-policy methods such as A3C~\cite{a3c}, PPO~\cite{ppo} or
TRPO~\cite{trpo}.

In all settings, we use the learning algorithm described in
Alg.~\ref{alg:dau} and supplementary, Alg.~1. The variants of DDPG and DQN
used are described in the supplementary material, as well as all the hyperparameters
used. We tested two different setups for DDPG and DQN.
In one, we scaled the discount factor (to avoid shortsightedness with small $\deltat$), but
left all other hyperparameters constant across time discretizations.
In the other, we used the properly rescaled discount
factor and reward from Eq.~\eqref{eq:def-gamma},
as well as $O(\deltat)$ learning rates for RMSProp.  The former yields slightly
better results, and is presented in the main text. Experiments on the second
variant are presented in the
supplementary material.

Let us stress that the quantities plotted are rescaled to make comparison
possible across different timesteps. For example,
return results are given in term of discretized return $R_\deltat$ as defined in \eqref{eq:discretized-return},\footnote{This mostly amounts to scaling rewards
by a factor $\deltat$ when this scaling is not naturally done in the environment. Environment-specific
details are given in the supplementary material.} and, most notably, time elapsed is always given in
physical time, i.e. the amount of time that the agent spent interacting with the environment (this is not the computational time).

\paragraph{Qualitative experiments: Visualizing policies and values.}
To provide qualitative results, and check robustness to time
discretization both in terms of raw return results and in terms
of convergence of the approximate value function and policies, we first provide results on the simple pendulum environment
from the OpenAI Gym classic control suite.  Its state space is of
dimension $2$. We visualize both the learnt value and policy functions by
plotting, for each point of the phase diagram $(\theta, \dot{\theta})$,
its value and policy. The results are presented in
Fig.~\ref{fig:pend} (value function) and in Fig.~1, 2 and 3 in
Supplementary.

>>>>

%! TEX root = icml_drau.tex
\section{Experiments}
\label{sec:exp}
\input{lc_fig.tex}
The experiments provided here are specifically aimed at showing that
the proposed method, DAU, works efficiently over a wide range of time
discretizations, without specific tuning, while standard deep $Q$-learning
approaches do not. DAU is compared to DDPG for continuous actions and to
DQN for
discrete actions.
Since we focus on Q-learning-based off-policy methods, we do not
compare against on-policy methods (A3C~\cite{a3c}, PPO~\cite{ppo},
TRPO~\cite{trpo}...).

In all setups, we use the algorithms described in
Alg.~\ref{alg:dau} and Supplementary Alg.~1. The variants of DDPG and DQN
used are described in the Supplementary, as well as all hyperparameters. We tested two different setups for DDPG and DQN.
In one, we scaled the discount factor (to avoid shortsightedness with small $\deltat$), but
left all other hyperparameters constant across time discretizations.
In the other, we used the properly rescaled discount
factor and reward from Eq.~\eqref{eq:def-gamma},
as well as $O(\deltat)$ learning rates for RMSProp.  The former yields slightly
better results, and is presented in the main text. Experiments on the second
variant are presented in the
supplementary material.

Let us stress that the quantities plotted are rescaled to make comparison
possible across different timesteps. For example,
returns are given in term of discretized return $R_\deltat$ as defined in \eqref{eq:discretized-return},\footnote{This mostly amounts to scaling rewards
by a factor $\deltat$ when this scaling is not naturally done in the environment. Environment-specific
details are given in the supplementary material.} and, most notably, time elapsed is always given in
physical time, i.e. the amount of time that the agent spent interacting with the environment (this is not the computational time).

\paragraph{Qualitative experiments: Visualizing policies and values.}
To provide qualitative results, and check robustness to time
discretization both in terms of returns and in terms
of convergence of the approximate value function and policies, we first provide results on the simple pendulum environment
from the OpenAI Gym classic control suite.  The state space is of
dimension $2$. We visualize both the learnt value and policy functions by
plotting, for each point of the phase diagram $(\theta, \dot{\theta})$,
its value and policy. The results are presented in
Fig.~\ref{fig:pend} (value function) and Figs.~1, 2,  3 in
Supplementary.

### 2019-01-23 paper/icml_drau/pendulum_fig.tex

%! TEX root = icml_drau.tex
\newcommand{\iw}{1cm}
\begin{figure*}[ht]
	\includegraphics[width=\textwidth]{figs_data/pendulum_unscaled/pendulum_fig_val_unscaled.pdf}
	\caption{Value functions obtained by DDPG (unscaled version) and AU at different instants in physical time of training on the pendulum swing-up environment. Each image represents the value function learnt, with $x$-axis representing angle, and $y$-axis angular velocity. The lighter the pixel, the higher the value.}
        \label{fig:pend}
\end{figure*}

>>>>

%! TEX root = icml_drau.tex
\newcommand{\iw}{1cm}
\begin{figure*}[ht]
	\includegraphics[width=\textwidth]{figs_data/pendulum_unscaled/pendulum_fig_val_unscaled.pdf}
	\caption{Value functions obtained by DDPG (unscaled version) and
	DAU at different instants in physical time of training on the
	pendulum swing-up environment. Each image represents the learnt
	value function (the $x$-axis is the angle, and the $y$-axis the angular velocity). The lighter the pixel, the higher the value.}
        \label{fig:pend}
\end{figure*}

### 2019-01-23 paper/icml_drau/supplementary.tex

\begin{proof}
	\input{policy_improvement.tex}
\end{proof}
\begin{theorem}
	Let $\cal A = \bb R^A$ be the action space, and let ${\cal P}_1 = \bb
	R^{p_1}$ and ${\cal P}_2 = \bb R^{p_2}$ be parameter spaces.  Let
	$A\colon {\cal P}_1 \times {\cal S} \times {\cal A} \rightarrow \bb R$
	and $V\colon {\cal P}_2 \times {\cal S} \rightarrow \bb R$ be ${\cal
	C}^2$ function approximators with bounded gradients and Hessians. Let
	$(a_t)_{t\geq 0}$ be a ${\cal C}^1$ exploratory actions trajectory and
	$(s_t)_{t\geq 0}$ the resulting state trajectory, when starting from $s_0$ and
	verifying Eq.~\eqref{eq:diff}.  Let $\theta^k_\deltat$ and
	$\psi^k_\deltat$ be the discrete parameter trajectories resulting from
	the gradient descent steps in the main text, with
	learning rates $\eta^V = \alpha^V \deltat^\beta$ and $\eta^A = \alpha^A \deltat^\beta$. Then,
	\begin{itemize}
		\item If $\beta = 1$ the discrete parameter trajectories converge to continuous parameter
			trajectories as $\deltat$ goes to $0$.
		\item If $\beta > 1$, parameter trajectories become
		stationary as
			$\deltat$ goes to $0$.
		\item If $\beta < 1$, parameters can grow arbitrarily large after an arbitrarily small physical time when $\deltat$ goes to $0$.
	\end{itemize}
\end{theorem}
\begin{proof}
	\input{proof_conv_params.tex}
	\input{no_conv_beta.tex}
\end{proof}

>>>>

\begin{proof}
	\input{policy_improvement.tex}
\end{proof}
\begin{theorem}
	Let $\cal A = \bb R^A$ be the action space, and let ${\cal P}_1 = \bb
	R^{p_1}$ and ${\cal P}_2 = \bb R^{p_2}$ be parameter spaces.  Let
	$A\colon {\cal P}_1 \times {\cal S} \times {\cal A} \rightarrow \bb R$
	and $V\colon {\cal P}_2 \times {\cal S} \rightarrow \bb R$ be ${\cal
	C}^2$ function approximators with bounded gradients and Hessians. Let
	$(a_t)_{t\geq 0}$ be a ${\cal C}^1$ exploratory action trajectory and
	$(s_t)_{t\geq 0}$ the resulting state trajectory, when starting from $s_0$ and
	following $ds_t/dt=F(s_t,a_t)$.  Let $\theta^k_\deltat$ and
	$\psi^k_\deltat$ be the discrete parameter trajectories resulting from
	the gradient descent steps in the main text, with
	learning rates $\eta^V = \alpha^V \deltat^\beta$ and $\eta^A =
	\alpha^A \deltat^\beta$ for some $\beta\geq 0$. Then,
	\begin{itemize}
		\item If $\beta = 1$ the discrete parameter trajectories converge to continuous parameter
			trajectories as $\deltat$ goes to $0$.
		\item If $\beta > 1$, parameter trajectories become
		stationary as
			$\deltat$ goes to $0$.
		\item If $\beta < 1$, parameters can grow arbitrarily large after an arbitrarily small physical time when $\deltat$ goes to $0$.
	\end{itemize}
\end{theorem}
\begin{proof}
	\input{proof_conv_params.tex}
	\input{no_conv_beta.tex}
\end{proof}
